Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2018-10-10 paper/paper.tex
Change layout to fit in 8pages

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.7\linewidth]{images/resources}
	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
	\label{resources}
\end{figure*}

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio. We chose Palladio to archive a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
This leads to the research question how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio. To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.55\linewidth]{images/process}
	\caption[]{The Simulation Process}
	\label{process}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{images/seff}
	\caption[]{Model of Job Behaviour}
	\label{seff}
\end{figure*}

@@

\begin{figure}
    \centering
    \begin{subfigure}{0.48\linewidth}
        \includegraphics[width=\linewidth]{images/validation/throughputs}
    \end{subfigure}
    \begin{subfigure}{0.48\linewidth}
        \includegraphics[width=\linewidth]{images/validation/utilizations}
    \end{subfigure}
    \caption{Measured and simulated throughputs and CPU efficiency, job data from May 2018}
    \label{fig:validation}
\end{figure}

To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types, and the overall CPU efficiency of the site.
These metrics describe how well the simulation matches the measured performance data.
Comparisons between  simulation results and measurements from the system itself for them are shown in figure \ref{fig:validation}.
The simulated throughput matches the reference with a difference of about 5.6\%.
The share of the simulated and measured job types also match accurately.

The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \ref{fig:validation}.
The initial ramp-up of the simulation is also visible.
The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of reproducing the metrics introduced above.
Runtimes of the simulated jobs were also measured, the average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.

Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
Thus, it would be feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.

>>>>

% layout changes
\setlength\belowcaptionskip{-15pt}
\renewcommand*{\bibfont}{\footnotesize}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.5\linewidth]{images/resources}
	\caption[]{Estimated CPU resources of ATLAS \cite{community}}
	\label{resources}
\end{figure*}

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio. We chose Palladio to archive a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
This leads to the research question how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio. To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Also, it can be used as a foundation for models and simulations of larger parts of the WLCG.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.45\linewidth]{images/process}
	\caption[]{The Simulation Process}
	\label{process}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.8\linewidth]{images/seff}
	\caption[]{Model of Job Behavior}
	\label{seff}
\end{figure*}

@@

\begin{figure}[h]
	\centering
	\vspace{-10px}
	\begin{subfigure}{0.48\linewidth}
		\includegraphics[width=\linewidth]{images/validation/throughputs}
	\end{subfigure}
	\begin{subfigure}{0.48\linewidth}
		\includegraphics[width=\linewidth]{images/validation/utilizations}
	\end{subfigure}
	\vspace{-10px}
	\caption{Measured and simulated throughputs and CPU efficiency, job data from May 2018}
	\label{fig:validation}
\end{figure}
\vspace{10px}
\noindent To validate the model and simulation we simulated CMS jobs at GridKa in May 2018.
The job performance data was split into two equal sized calibration and test datasets by randomly sampling reports from the full data set.
For the validation we considered the following metrics: Overall throughput of jobs, shares of the job types, and the overall CPU efficiency of the site.
These metrics describe how well the simulation matches the measured performance data.
Comparisons between  simulation results and measurements from the system itself for them are shown in figure \Cref{fig:validation}.
The simulated throughput matches the reference with a difference of about 5.6\%.
The share of the simulated and measured job types also match accurately.

The CPU utilization in the simulation and a comparison with the measured CPU efficiency is also shown in figure \Cref{fig:validation}.
The initial ramp-up of the simulation is also visible.
The overall average utilization in the simulation deviates from the reference by about 3\%, showing that even the current simplified model is capable of reproducing the metrics introduced above.
Runtimes of the simulated jobs were also measured, the average runtime of simulated jobs deviates by approximately 8\% from the measured runtimes in the test data set.

Simulating one week of CMS jobs at GridKa takes about 24 minutes on a laptop.
Thus, it is feasible to simulate several sites at once in a reasonable time, however simulating larger parts of the grid is future work.
